---
title: 'Using `r icons::fontawesome("git-square")` with `r icons::fontawesome("r-project")`'
#subtitle: "slides : [https://tinyurl.com/2s4antdy](https://tinyurl.com/2s4antdy)"
author: "Maxime Jaunatre"
institute: "LESSEM | INRAE" 
date: "2022-12-01 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, metropolis, metropolis-fonts]
    nature:
      highlightStyle: github
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
      seal: false
---

```{r setup, include=FALSE}
library(knitr)
library(emo)
library(icons)
options(htmltools.dir.version = FALSE, knitr.duplicate.label = 'allow')
knitr::opts_chunk$set(fig.align = "center", eval = FALSE)
```

```{r, eval = FALSE, include = FALSE}
# This are the dependencies to compile this rmd.
remotes::install_github("mitchelloharawild/icons")
icons::download_fontawesome()
remotes::install_github("hadley/emo")
```


exclude: true

class: inverse, center, middle

# Introduction

---
exclude: true

# Utilisation de Git dans un travail avec R

**Résumé :** 

---

class: inverse, center, middle

# Remote repository : Github


---
# Working with others

```{r bordel, engine='tikz', fig.ext = 'png', engine.opts = list(template = "tex/gitdags.tex"), echo = FALSE, eval=TRUE, cache= TRUE}
\begin{tikzpicture}
  \matrix (m)
    [
      matrix of nodes,
      nodes in empty cells,
      column sep      = 3em,
      row sep         = 5ex,
      column 1/.style = { nodes = { empty } },
      column 2/.style = { nodes = { empty } },
      column 3/.style = { nodes = { empty } },
      column 4/.style = { nodes = { empty } },
    ]
    {
        &  &  &  \\
        % A.txt & A.txt & A.txt &  \\
        A1.txt & A1'.txt & A1''.txt &  \\
        A2.txt & A2'.txt &  & B.txt \\
        A2.txt &  & B1.txt & B1'.txt \\
        & C.txt & B2.txt &  \\
        & D.txt &  &  \\
    };
    \node[fit=(m-1-1)(m-1-1)]{Etno P.};
    \node[fit=(m-1-2)(m-1-2)]{Gorgious K.};
    \node[fit=(m-1-3)(m-1-3)]{Candy H. C.};
    \node[fit=(m-1-4)(m-1-4)]{Bud B.};
  
    \draw [arrow] (m-2-1) -- (m-3-1);
    \draw [arrow] (m-3-1) -- (m-4-1);
    \draw [arrow] (m-4-1) -- (m-5-2);

    \draw [arrow] (m-2-2) -- (m-3-2);
    \draw [arrow] (m-3-2) -- (m-4-1);
    \draw [arrow] (m-5-2) -- (m-6-2);

    \draw [arrow] (m-2-3) -- (m-3-4);
    \draw [arrow] (m-4-3) -- (m-5-3);
    \draw [arrow] (m-4-3) -- (m-5-2);
    \draw [arrow] (m-5-3) -- (m-6-2);
    
    \draw [arrow] (m-3-4) -- (m-4-4);
    \draw [arrow] (m-3-4) -- (m-4-3);
    \draw [arrow] (m-4-4) -- (m-5-3);

\end{tikzpicture}
```

It becomes complex very quickly... 12 versions of a single file?

---
# Using a remote repository

Git is installed on your computer, and a repository in your folder.

Principle of **Cloud** : installation of Git on a server and copy of the repository.

Several hosters exist:

**Github** (Microsoft) `r icons::fontawesome("github")`   
**Bitbucket** `r icons::fontawesome("bitbucket")`   
**Gitlab** `r icons::fontawesome("gitlab")`

And the list is long: [https://git.wiki.kernel.org/index.php/GitHosting](https://git.wiki.kernel.org/index.php/GitHosting)

It is possible to copy a repository on several hosts (mirroring) !

*You can even have a remote on the same computer, or on an external disk ! For example another disk...but the reality is a bit more complex.*

---
# Gorgious's repository

**Repo** : github.com/Gorgious/name

```{r github, engine='tikz', fig.ext = 'png', engine.opts = list(template = "tex/gitdags.tex"), echo = FALSE, eval=TRUE, cache= TRUE}

\begin{tikzpicture}
  \matrix (m)
    [
      matrix of nodes,
      nodes in empty cells,
      column sep      = 3em,
      row sep         = 5ex,
      column 1/.style = { nodes = { empty } },
      column 2/.style = { nodes = { empty } },
      column 3/.style = { nodes = { empty } },
    ]
    {
        &  &  \\
        & 7f21c09 & \\  
        & d39ac60 & \\  
        & c78c64d & \\  
        &  & \\ 
    };
     \node[fit=(m-1-1)(m-1-1)]{Collaborator};
     \node[fit=(m-1-2)(m-1-2)]{Github};
     \node[fit=(m-1-3)(m-1-3)]{Local};
  
    \foreach \i/\j in {2/3, 3/4} {
      \draw [arrow] (m-\i-2) -- (m-\j-2);
    }
\end{tikzpicture}
```

---
# Clone a remote

```{r clone, engine='tikz', fig.ext = 'png', engine.opts = list(template = "tex/gitdags.tex"), echo = FALSE, eval=TRUE, cache= TRUE}

\begin{tikzpicture}
  \matrix (m)
    [
      matrix of nodes,
      nodes in empty cells,
      column sep      = 3em,
      row sep         = 5ex,
      column 1/.style = { nodes = { empty } },
      column 2/.style = { nodes = { empty } },
      column 3/.style = { nodes = { empty } },
    ]
    {
        &  &  \\
        & 7f21c09 & 7f21c09 \\  
        & d39ac60 & d39ac60 \\  
        & c78c64d & c78c64d \\  
        &  &  \\  
    };
     \node[fit=(m-1-1)(m-1-1)]{Collaborator};
     \node[fit=(m-1-2)(m-1-2)]{Github};
     \node[fit=(m-1-3)(m-1-3)]{Local};
  
    \foreach \i/\j in {2/3, 3/4} {
      \draw [arrow] (m-\i-2) -- (m-\j-2);
      \draw [arrow] (m-\i-3) -- (m-\j-3);
    }
\end{tikzpicture}
```

---
# Local commit

```{r local, engine='tikz', fig.ext = 'png', engine.opts = list(template = "tex/gitdags.tex"), echo = FALSE, eval=TRUE, cache= TRUE}

\begin{tikzpicture}
  \matrix (m)
    [
      matrix of nodes,
      nodes in empty cells,
      column sep      = 3em,
      row sep         = 5ex,
      column 1/.style = { nodes = { empty } },
      column 2/.style = { nodes = { empty } },
      column 3/.style = { nodes = { empty } },
    ]
    {
        &  &  \\
        & 7f21c09 & 7f21c09 \\  
        & d39ac60 & d39ac60 \\  
        & c78c64d & c78c64d \\  
        &  & |[high]| cbdac1e \\  
    };
     \node[fit=(m-1-1)(m-1-1)]{Collaborator};
     \node[fit=(m-1-2)(m-1-2)]{Github};
     \node[fit=(m-1-3)(m-1-3)]{Local};
  
    \foreach \i/\j in {2/3, 3/4} {
      \draw [arrow] (m-\i-2) -- (m-\j-2);
      \draw [arrow] (m-\i-3) -- (m-\j-3);
    }
  \draw [harrow] (m-4-3) -- (m-5-3);
\end{tikzpicture}
```

---
# Push

```{r push, engine='tikz', fig.ext = 'png', engine.opts = list(template = "tex/gitdags.tex"), echo = FALSE, eval=TRUE, cache= TRUE}

\begin{tikzpicture}
  \matrix (m)
    [
      matrix of nodes,
      nodes in empty cells,
      column sep      = 3em,
      row sep         = 5ex,
      column 1/.style = { nodes = { empty } },
      column 2/.style = { nodes = { empty } },
      column 3/.style = { nodes = { empty } },
    ]
    {
        &  &  \\
        & 7f21c09 & 7f21c09 \\  
        & d39ac60 & d39ac60 \\  
        & c78c64d & c78c64d \\  
        & |[high]| cbdac1e & |[high]| cbdac1e \\  
    };
     \node[fit=(m-1-1)(m-1-1)]{Collaborator};
     \node[fit=(m-1-2)(m-1-2)]{Github};
     \node[fit=(m-1-3)(m-1-3)]{Local};
  
    \foreach \i/\j in {2/3, 3/4} {
      \draw [arrow] (m-\i-2) -- (m-\j-2);
      \draw [arrow] (m-\i-3) -- (m-\j-3);
    }
    \draw [harrow] (m-4-2) -- (m-5-2);
    \draw [harrow] (m-4-3) -- (m-5-3);
\end{tikzpicture}
```

---
# Late

```{r retard, engine='tikz', fig.ext = 'png', engine.opts = list(template = "tex/gitdags.tex"), echo = FALSE, eval=TRUE, cache=TRUE}

\begin{tikzpicture}
  \matrix (m)
    [
      matrix of nodes,
      nodes in empty cells,
      column sep      = 3em,
      row sep         = 5ex,
      column 1/.style = { nodes = { empty } },
      column 2/.style = { nodes = { empty } },
      column 3/.style = { nodes = { empty } },
    ]
    {
        &  &  \\
        7f21c09 & 7f21c09 & 7f21c09 \\  
        d39ac60 & d39ac60 & d39ac60 \\  
        c78c64d & c78c64d & c78c64d \\  
         & cbdac1e & cbdac1e \\
    };
     \node[fit=(m-1-1)(m-1-1)]{Collaborator};
     \node[fit=(m-1-2)(m-1-2)]{Github};
     \node[fit=(m-1-3)(m-1-3)]{Local};
  
    \foreach \i/\j in {2/3, 3/4} {
      \draw [arrow] (m-\i-1) -- (m-\j-1);
      \draw [arrow] (m-\i-2) -- (m-\j-2);
      \draw [arrow] (m-\i-3) -- (m-\j-3);
    }
    \draw [arrow] (m-4-2) -- (m-5-2);
    \draw [arrow] (m-4-3) -- (m-5-3);
\end{tikzpicture}
```

---
# Pull

```{r pull, engine='tikz', fig.ext = 'png', engine.opts = list(template = "tex/gitdags.tex"), echo = FALSE, eval=TRUE, cache = TRUE}

\begin{tikzpicture}
  \matrix (m)
    [
      matrix of nodes,
      nodes in empty cells,
      column sep      = 3em,
      row sep         = 5ex,
      column 1/.style = { nodes = { empty } },
      column 2/.style = { nodes = { empty } },
      column 3/.style = { nodes = { empty } },
    ]
    {
        &  &  \\
        7f21c09 & 7f21c09 & 7f21c09 \\  
        d39ac60 & d39ac60 & d39ac60 \\  
        c78c64d & c78c64d & c78c64d \\  
        |[high]| cbdac1e & cbdac1e & cbdac1e \\
    };
     \node[fit=(m-1-1)(m-1-1)]{Collaborator};
     \node[fit=(m-1-2)(m-1-2)]{Github};
     \node[fit=(m-1-3)(m-1-3)]{Local};
  
    \foreach \i/\j in {2/3, 3/4} {
      \draw [arrow] (m-\i-1) -- (m-\j-1);
      \draw [arrow] (m-\i-2) -- (m-\j-2);
      \draw [arrow] (m-\i-3) -- (m-\j-3);
    }
    \draw [arrow] (m-4-2) -- (m-5-2);
    \draw [arrow] (m-4-3) -- (m-5-3);
    \draw [harrow] (m-4-1) -- (m-5-1);
\end{tikzpicture}
```


---
# COMMIT -> PULL -> PUSH

```{r pull_push_diag, engine='tikz', fig.ext = 'png', engine.opts = list(template = "tex/gitdags.tex"), echo = FALSE, eval=TRUE, cache = TRUE}
\begin{tikzpicture}[block/.style={draw,fill=strange,minimum height=2.5em},
    font=\sffamily,>=stealth]

    \begin{scope}[start chain=A going below,node distance=1em,
                  local bounding box=buffers]
        \node[block,minimum width=6em](remote){Remote github};
        \node[right= 8em of remote, block,minimum width=6em](local){Fichiers};
    \end{scope} 

    \node[below= 6em of buffers,align=center, block,minimum width=6em](repo){Repository \\Git local};

    %\draw[->, thick](remote) edge[bend left] node[midway, fill=white]{Pull} (local);
    \draw[->, thick](remote) edge[bend left] node[midway, fill=white]{Pull} (repo);
    %\draw[->, thick](remote) edge[dashed] node[midway, fill=white]{Fetch} (repo);
    \draw[->, thick](remote) edge node[midway, fill=white]{Clone} (local);
    \draw[->,thick](local) edge[bend left] node[midway, fill=white]{Add} (repo);
    \draw[->,thick](repo) edge[bend left] node[midway, fill=white]{Push} (remote);
    \draw[->,thick](repo) edge[loop right] node[midway, fill=white]{Commit} (repo);
    

\end{tikzpicture}
```
